---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# twiGPS
## Time-Weighted spatial averaging (TWSA) methods Integrated with GPS measured environmental exposure
<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/Graczykowski/twsagps/branch/master/graph/badge.svg)](https://app.codecov.io/gh/Graczykowski/twsagps?branch=master)
[![R-CMD-check](https://github.com/Graczykowski/twsagps/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Graczykowski/twsagps/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`twiGPS` is a R package designed for calculating exposure and activity space from GPS data. You can utilize 
Time-Weighted Spatial Averaging (TWSA) methods to recognize distinct approaches to representing spatially exposure and compare these methods.


## Installation

You can install the development version of twiGPS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Graczykowski/twiGPS")
```

## Usage

You are able to choose from 4 distinct TWSA methods by choosing those package functions: `exposure_PO()` (Point Overlay), `exposure_KDE()` (Kernel Density Estimation), `exposure_DR()` (Density Ranking) and `exposure_LS()` (Line Segment).

You can start from data.frame with GPS coordinates:

```{r example, message=FALSE, warning=FALSE}
library(twiGPS)
library(terra)

KDE = exposure_KDE(geolife_sandiego, coords = c("lon", "lat"), cellsize = 50,
                   bandwidth = 200, input_crs = "EPSG:4326", output_crs = "EPSG:32611")
plot(KDE)
```
or you can use sf and terra SpatVector objects:
```{r example_2, warning=FALSE, message=FALSE}
geolife_v = vect(geolife_sandiego, geom = c("lon", "lat"), crs = "EPSG:4326")
geolife_v = project(geolife_v, "EPSG:32611")

LS = exposure_LS(geolife_v, cellsize = 50, bandwidth = 200)
plot(LS)
```
When you want to implement your environmental data to calculate environmental exposure instead of activity space you should use `env_data` argument:
```{r example_3, warning=FALSE, message=FALSE}
ndvi = rast(system.file("extdata/landsat_ndvi.tif", package = "twiGPS"))
DR_env = exposure_DR(geolife_v, bandwidth = 200, env_data = ndvi)
plot(DR_env)
```
You can compare statistics of all TWSA methods using `exposure_stats()`:
```{r example_4, message=FALSE, warning=FALSE}
statistics = c("area", "min", "max", "mean", "std", 'sum')

exposure_stats(KDE, LS, DR_env, stats = statistics,
               row_names = c("KDE", "Line Segment",
                             "Density Ranking NDVI"))
```
